This is an example of article title

### 2025-04-10 paper/paper.tex
pushing updates preparing for final review

\section{Introduction}
\label{sec:intro}
Cast parts may have small geometric variations from lot to lot that need to be addressed before machining by altering the CNC code.
Current practice is dominated by iterative adjustments by the human operator, which requires highly trained workers, takes a long time, and still, may produce scrap.
%TODO: kell egyáltalán ezeket hivatkozni?
Automated methods exist for complex free-form parts like wind turbines that place the entire blank as a single solid object \cite{ding:2021_CoarsefineOptimizationMethod}\cite{tan:2014_UnconstrainedApproachBlank}.
Multi-operation blank localization~\cite{cserteg:2023_Annals} however handles group of features independently providing greater flexibility than traditional approaches.
It focuses on drilling and milling which are among the most common machining operations, making it applicable to a wide range of products.
The abstract method and its implementation were developed in parallel, which required a language with support for easy prototyping and wide variety of tools.
Exactly for these reasons we chose the Julia language~\cite{bezanson2017julia}.

@@

The problem involves looking for the optimal position for each to-be-machined feature (short: machined feature) on the workpiece.
Based on the actual machining operations, these features are grouped into feature groups, for example, drilled holes on the same side of the part, whose drilling positions are defined relative to a common reference (the part zero).
Features within the same group move together, but different feature groups can move separately.
The machined features need to "enclose" the pre-cast features on the blank (called rough features) with a minimum machining allowance which ensures the required surface finish.
While doing so, the dimensional tolerances between some features need to be respected, as these describe functional properties (e.g. connections to mating parts).
Fig.~\ref{fig:hatfig} from \cite{cserteg:2023_Annals} shows an example of these features.

\begin{figure}[hbt]
	\centerline{\includegraphics[width=0.95\columnwidth]{cirp-annals-2023-figure-2.png}}
	\caption{3D scanned rough features (in grey), two machined feature groups (orange and purple) with their part zeros, and tolerances connecting features \cite{cserteg:2023_Annals}.}
	%\caption{Measured part with two machined feature groups and tolerances between them. $r^M$ and $r^R$ denote the machined and rough radii of a feature, while $\delta$ the machining allowance \cite{cserteg:2023_Annals}.}
	\label{fig:hatfig}
\end{figure}

Following common machining practice the objective of the optimization program is to achieve as little tolerance deviation as possible, while ensuring that material is removed for proper surface finish.
The features are moved through their part zeros, which are the variables of the developed convex quadratically constrained quadratic program (convex QCQP).

A tolerance model was developed to replicate the tolerances used in design and machining process.
This model encodes the distance of machined-machined (or sometimes machined-rough) features as axis-aligned minimum and maximum distance.
These distances form constraints in the optimization program, while minimizing the deviation relative to their mean (central?) value across all tolerances is the objective function itself.

The allowance calculation requires the final part specification in the form of the machining CNC code as well as a representation of the rough geometries.
From these input data the rough-machined feature geometry pairs need to be extracted so that their machining allowance can be computed.
The geometric parameters are extracted so the constraint equations for the minimum machining allowance can be generated for the optimization program.

The optimization model itself and use-cases are described in \cite{cserteg:2023_CMS} and \cite{cserteg:2023_Annals}, while implementation details are given in the following section.

@@

%Evaluating the minimum machining allowance contributes to the constraints of the optimization program and depends on geometric calculations as well.
%The allowance calculation requires the position of the features as well as other geometric parameter (radii of cylinders for example).
%TODO: még egy mondat?

@@

During the development of the algorithm, we needed a software tool that supports the quick prototyping needs of research, while giving a solid foundation to validate the concept with our industry partner.
We needed a tool that:

\begin{itemize}
	\item Interfaces our model to the optimization solver and enables rapid prototyping.
	\item Supports a variety of geometrical representations, especially regarding the differences of drilling/milling operations and free-form/primitive geometry representation.
	\item Supports importing geometries from a variety of measurement tools (CMM, 3D scanner, measurement arm).
%	\item Supports easy extension of the geometries if required.
%	\item Supports extending 
%	\item Translates the tolerance model for the optimization solver.
	\item Supports easy debugging and helps analyzing and visualizing the results.
\end{itemize}

The heart of the implementation is the solver provided by the JuMP ecosystem~\cite{Lubin2023}.
With JuMP's excellent design, implementing the optimization model was as easy as repeating the mathematical model in Julia.
Other advantage of the JuMP ecosystem is that solvers can be easily swapped if needed.
For development, the FICO Xpress solver was used, but our industry partner could use the Ipopt or SCIP solvers without issue.

@@

To generate the optimization model we needed to store the 1. nc code 2. rough geometry 3. tolerances.
A common thing regarding the first two, that some kind of geometric representation is needed.
Representing the NC code can be done with plane and cylinder geometries for milling and drilling operations.
While the rough geometry poses greater challanges.
Different measurements instruments output different type of geometric data, for example a coordinate measurement arm (CMM) will provide primitive geometry definitions like disks and cylinders while a 3D scanner output point clouds or meshes (pongyolán csak free form).
Property of the optimization model is that it is built around "feature points" (consistently defined points of different geometries).
These two aspects required us to develop a type tree.
To cover our current needs (drilling and milling operations), planar and cylindrical geometries are defined.
The other property of geometries is if they are primitive or free form, which is covered by the holy trait pattern.
\texttt{IsPrimitve} or \texttt{IsFreeForm} traits can be applied to geometry types independently of their planar or cylindrical type.
Code block~\ref{lst:def-types} shows the core of the implementation of the type system.

\begin{lstlisting}[language = Julia, numbers=left, label={lst:def-types}, caption={Draft of the type system used by \texttt{BlankLocalizationCore.jl}.}]
# Supertype for localization geometries.
# "A": abbr. for Abstract
# "L": abbr. for Localization
abstract type ALocGeometry end
abstract type AHoleGeometry <: ALocGeometry end
abstract type APlaneGeometry <: ALocGeometry end

@@

The other part of the optimization generation interface is definition of the query functions that work on these types.
Free form geometries need to implement the \texttt{surfacepoints} and \texttt{filteredsurfacepoints} methods;
Primitive geometries need to have the \texttt{featurepoint} method and cylindrical features need the \texttt{featureradius} on top of that.
For visualization the \texttt{visualizationgeometry} function need to be defined whose output is passed to \texttt{Meshes.viz} function, so Meshes objects are needed.
A short example for such a definition is shown in code block~\ref{lst:example}.

The visulization interface is built around Meshes.jl, therefore the to-be-showed geometries must be from that package.
Using the Meshes ecosystem, we could produce publication quality images, like Fig.~\ref{fig:hatfig} from \cite{cserteg:2023_Annals}.

\begin{lstlisting}[language = Julia, numbers=left, label={lst:example}, caption={Defining a new type.}]
struct MyDisk <: AHoleGeometry
  p::Vector{Float64} # center point
  n::Vector{Float64} # surface normal
  d::Float64 # diameter
end
GeometryStyle(::Type{MyDisk}) = IsPrimitive()

@@

The optimization program uses a feature point (or feature points for free form geometries) of features, that is used for computing the machining allowance and the dimensional tolerances.
A function interface is defined for querying these feature points of geometries, that is used to generate the optimization program.
Code block~\ref{lst:example} shows the definition of those functions as well.
The visulization interface is built around Meshes.jl, therefore the to-be-showed geometries must be from that package.
Using the Meshes ecosystem, we could produce publication quality images, like Fig.~\ref{fig:hatfig} from \cite{cserteg:2023_Annals}.

@@

As mentioned earlier, an important requirement was to handle the output formats of many different devices.
Every measurement instrument comes with its own processing software and export formats, not necessarily designed for interoperability.
As a result, we needed to write (or use) parsers for different text and tabular formats.

>>>>

\section{Introduction}
\label{sec:intro}
Cast parts may have small geometric variations from lot to lot that need to be addressed before machining by altering the CNC code.
Current practice is dominated by iterative adjustments by human operators, which requires highly trained workers and takes a long time.
%, and still, may produce scrap.
%TODO: kell egyáltalán ezeket hivatkozni?
Automated methods exist for complex free-form parts like wind turbines that place the entire blank as a single solid object \cite{ding:2021_CoarsefineOptimizationMethod}\cite{tan:2014_UnconstrainedApproachBlank}.
Multi-operation blank localization~\cite{cserteg:2023_Annals} however handles group of features independently providing greater flexibility than traditional approaches.
It focuses on drilling and milling which are among the most common machining operations, making it applicable to a wide range of products.
The abstract method and its implementation were developed in parallel, which required a language with wide variety of tools and support for easy prototyping.
Exactly for these reasons we chose the Julia language~\cite{bezanson2017julia}.

@@

The problem involves looking for the optimal position for each to-be-machined feature (short: machined feature) on the workpiece.
%chatgpt
These features are grouped together, with each group defined relative to a specific reference point known as the part zero.
The layout of these part zeros is controlled by the structure of the machining CNC code.
By moving the part zeros, we can indirectly control the positions of the associated features.
Each part zero—and by extension, each group—can be moved independently, which gives the method its flexibility.
In the optimization program, the positions of the part zeros are the decision variables.

%eredeti
\iffalse
These features are structured into groups, each group being defined relative to a common reference, called part zero.
This is defined by the structure of the machining CNC code.
This means that we can influence the features' positions through moving the part zeros.
Part zeros (thus the groups) can be moved separately, which gives the method its flexibility.
The position of the part zeros are the decision variables of the optimization program.
\fi
%Features within the same group must be moved together, but different feature groups can be moved separately.
%TODO: optimalizáló leírás:
% 0. motiváció
% 1. mik a döntések
% 2. mik a korlátok
% 3. mi a célfv

The machined features need to "enclose" the pre-cast features on the blank (called rough features) with a minimum machining allowance which ensures the required surface finish.
The allowance calculation requires the final part specification in the form of the machining CNC code as well as a representation of the rough geometries.
From their positions and geometric parameters their distance can be computed, which then can be used to generate the allowance constraints for the optimization program.

The other set of constraints roots from respecting the dimensional tolerances describing functional properties (e.g. connections to mating parts).
The developed tolerance model encodes the distance of machined-machined (or sometimes machined-rough) features as axis-aligned minimum and maximum distance.
%These distances form constraints in the optimization program, while minimizing the deviation relative to their mean (central?) value across all tolerances is the objective function itself.
Fig.~\ref{fig:hatfig} from \cite{cserteg:2023_Annals} shows some examples for these features.

\begin{figure}[b]
	\centerline{\includegraphics[width=0.95\columnwidth]{cirp-annals-2023-figure-2.png}}
	\caption{3D scanned rough features (in grey), two machined feature groups (orange and purple) with their part zeros, and tolerances connecting the machined features~\cite{cserteg:2023_Annals}.}
	%\caption{Measured part with two machined feature groups and tolerances between them. $r^M$ and $r^R$ denote the machined and rough radii of a feature, while $\delta$ the machining allowance \cite{cserteg:2023_Annals}.}
	\label{fig:hatfig}
\end{figure}

%A tolerance model was developed to replicate the tolerances used in design and machining process.

% vagy full stop, vagy "while ensuring the above constraints are respected
Following common machining practice, the objective of the optimization program is to achieve as little tolerance deviation as possible.
The problem is formulated as a convex quadratically constrained quadratic program (convex QCQP).
%, where decision variables are the positions of the part zeros.
%TODO az hogy a part zeroval mozgatunk, az legyen tisztázva korábban
%The features are moved through their part zeros, which are the variables of the developed .
The optimization model itself and use-cases are described in \cite{cserteg:2023_CMS} and \cite{cserteg:2023_Annals}, while implementation details are given in the following section.

@@

%Evaluating the minimum machining allowance contributes to the constraints of the optimization program and depends on geometric calculations as well.
%The allowance calculation requires the position of the features as well as other geometric parameter (radii of cylinders for example).

@@

%The following aspects of julia enabled us to develop the model and its implementation quickly and efficiently
%During the development of the algorithm, the following requirements were formulated::
%we needed a software tool that supports the quick prototyping needs of research, while giving a solid foundation to validate the concept with our industry partner.
%Summarizin
%Julia helped us to develop the implementation that had the following properties

%The implementation needed to be ilyen, which 
Julia enabled us to develop an implementation with the following properties:
\begin{itemize}
	%TODO: vagy legyen két rész a rapid prototypinggal vagy skip
	\item Concise interface to generate the parameters of the declarative optimization program based on procedural geometric calculations.
%	\item Interfaces our model to the optimization solver and enables rapid prototyping.
	\item Support for a variety of geometrical representations, especially regarding the differences of drilling/milling operations and free-form/primitive geometry representations.
	\item Support for analyzing and visualizing the results.
\end{itemize}

%TODO: asbtract type system is a good base for handling the import fnctions for the measuremnt tool

The heart of the implementation is the solver provided by the JuMP ecosystem~\cite{Lubin2023}.
With JuMP's excellent design, combining the necessary geometric calculations and the declarative optimization program definition was straightforward.
%implementing the optimization model was as easy as repeating the mathematical model in Julia.
%TODO
%jjumps support for declarative programming meets with julia's abstract type system and procedural techniques for geometrical computations
% az együtthatókat könnyű volt juliaból generálni, mert absztrakt módon könnyű megkeresni őket (vagy mi)
Other advantage of JuMP is that solvers can be easily swapped.
For development, the (commercial) FICO Xpress solver was used, but our industry partner could use the (open source) Ipopt or SCIP solvers without issue.

%TODO: absztrakt typusokra deinfinájuk a funckiókat, amiket aztán alacsony sznten könnyű implementálni

@@

To handle the tolerance and allowance calculations, the part definition (CNC code), rough geometry measurements and tolerances need to be stored.
Their representation is built upon a flexible geometry type system.
%To generate the optimization model we needed to store the 1. nc code 2. rough geometry 3. tolerances.
%A common thing regarding the first two, that some kind of geometric representation is needed.

The CNC code can be represented with plane and cylinder geometries for milling and drilling operations.
This "type" information of the geometries is encoded with Julia types.
The primitive or free-form nature of the geometries is implemented with the holy traits pattern and is necessary because different instruments output different types of geometric data.
For example, a coordinate measurement machine will provide primitive geometry definitions like disks and cylinders, while a 3D scanner outputs point clouds or meshes (called free-form collectively).
The \texttt{IsPrimitve} or \texttt{IsFreeForm} traits are applied to geometry types independently of their planar or cylindrical type.
Code block~\ref{lst:def-types} shows a shortened version of the implemented type system.

\begin{lstlisting}[language = Julia, numbers=left, label={lst:def-types}, caption={Shortened implementation of the type system used by \texttt{BlankLocalizationCore.jl}.}]
# Type tree for localization geometries.
# "ALoc": AbstractLocalization
abstract type ALocGeometry end
abstract type AHoleGeometry <: ALocGeometry end
abstract type APlaneGeometry <: ALocGeometry end

@@

The optimization model uses a "feature point" concept which requires the position and some parameters of the geometries.
A function interface is designed for accessing these values.
%is that it is built around "feature points" (consistently defined points of different geometries).
%The other part of the optimization generation interface is definition of the query functions that work on these types.
The package documentation contains the list of functions that need to be defined but some are showcased in code block~\ref{lst:example}.
%Free form geometries need to implement the \texttt{surfacepoints} and \texttt{filteredsurfacepoints} methods;
%Primitive geometries need to have the \texttt{featurepoint} method and cylindrical features need the \texttt{featureradius} on top of that.
One function that we want to highlight is the \texttt{visualizationgeometry}, which needs to return a \texttt{Meshes.jl} object that will be passed to the \texttt{Meshes.viz} function.
%The visulization interface is built around Meshes.jl, therefore the to-be-showed geometries must be from that package.
Using the Meshes ecosystem~\cite{Hoffimann2023}, we could not only interactively inspect the results of the optimization, but also produce publication quality images, like Fig.~\ref{fig:hatfig} in \cite{cserteg:2023_Annals}.

\vspace*{2em}

\begin{lstlisting}[language = Julia, numbers=left, label={lst:example}, caption={Defining a new type for the optimization model.}]
struct MyDisk <: AHoleGeometry
  p::Vector{Float64} # center point
  n::Vector{Float64} # surface normal
  d::Float64 # diameter
end

GeometryStyle(::Type{MyDisk}) = IsPrimitive()

@@

\iffalse
The optimization program uses a feature point (or feature points for free-form geometries) of features, that is used for computing the machining allowance and the dimensional tolerances.
A function interface is defined for querying these feature points of geometries, that is used to generate the optimization program.
Code block~\ref{lst:example} shows the definition of those functions as well.
The visulization interface is built around Meshes.jl, therefore the to-be-showed geometries must be from that package.
Using the Meshes ecosystem, we could produce publication quality images, like Fig.~\ref{fig:hatfig} from \cite{cserteg:2023_Annals}.

@@

As mentioned earlier, an important requirement was to handle the output formats of many different devices.
Every measurement instrument comes with its own processing software and export formats, not necessarily designed for interoperability.
As a result, we needed to write (or use) parsers for different text and tabular formats.
\fi
